---
title: "Wine quality"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Background
In this dataset, several parameters of different bottles of wine were recorded. These bottles were rated by professional wine experts. They rated the `quality` of a wine on an ordinal scale: 1 - poor, 2 - average, 3 - excellent.


# Task


1) Standardize variables
You need to standardize your continuous predictors.

2) Build models
You need to define and calculate an ordinal regression model. This model should contain every predictor in the dataset. Also create a model where `density` is not used, one where `pH` is not used, and one where neither `density` nor `pH` are used.

3) Compare models
Create tables that contain all parameter estimates, std. errors, t statistics, and p values.
Also create a table that contains model diagnostics, and choose the best model based on AIC.



# Solution
```{r setup}
library(tidyverse)
library(broom)
library(performance)
library(ordinal)
library(sjPlot)

wine_data <- read_csv(here::here("data/wine_quality.csv"))
```

# Standardize continuous variables

```{r}

```

# Building the models
```{r}

```

# Compare models

```{r}

```

# Printing the results
```{r}

```