---
title: "Risk aversion"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Background
In this study, we investigated how age and religion influences risk aversion. Studies show that older adults make safer choices than younger adults and religious people people also make safer choices than non religious people. We do not assume an interaction between age and religiosity. We will try to confirm these previous findings in this study.

In the study, the participants had to complete a two-choice risky decision making paradigm, a gambling task. The two choices that were presented to the participants had different risk-gain proportions. One choice was always more risky than the other with higher benefit.

We measured religiosity with a 10 point Likert-type scale.

In this task, you will have to test the main hypothesis a generalized linear regression.

## Codebook
`codebooks/riskeversion_codebook.txt`

# Task
1) Build models
You need to build generalized linear regression model testing our hypothesis.

2) Report results
Create a single table that contains all parameter estimates, std. errors, t statistics, and p values.

# Solution
```{r setup}
library(tidyverse)
risk_data <- read_csv(here::here("data/riskaversion_data.csv"))
```

# Building the models
```{r}
# Creating a dataset that contains the result of the model in a tidy way
# Data manipulation
risk_models <- 
  risk_data %>% 
  mutate(
    choice = as.factor(choice),
    choice = fct_relevel(choice, "safe", "risky")
  ) %>%
  pivot_longer(cols = religion:age, names_to = "predictor_name", values_to = "predictor_value") %>% 
  group_by(predictor_name) %>% 
  nest() %>% 
  mutate(
    model = map(data,
                ~ glm(choice ~ predictor_value, family = "binomial", data = .x) %>% 
                  broom::tidy()
                )
    )
```

# Printing the results in a table
```{r}
# Printing results using the gt package
library(gt) 

risk_models %>% 
  select(-data) %>% 
  unnest(model) %>% 
  # Rename the predictor, so model terms will be more specific
  mutate(term = str_replace(term, "predictor_value", predictor_name)) %>%
  gt() %>% 
  # Formatting numers to be 2 digits, except for p value (3 digits)
  fmt_number(c("estimate", "std.error", "statistic"), decimals = 2) %>% 
  fmt_number("p.value", decimals = 3) %>% 
  tab_options(row_group.background.color = "grey50")
```

