---
title: "Doctor visits"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Background
In this study, we will use cross-section data from an Australian health survey between 1977 and 1978. Your task is to predict the number of illnesses in the past two weeks based on the self-reported health score, while controlling fo the income, age, and gender.

## Codebook
```{r}
?AER::DoctorVisits
```

# Task
1) Build models
You need to build Poisson regression model for testing the hypothesis.

2) Report results
Create a single table that contains all parameter estimates, std. errors, t statistics, and p values.

3) Check assumptions and adjust analysis if necessary.

# Solution
```{r setup}
library(tidyverse)
library(performance)
data("DoctorVisits", package = "AER")
```

# Building the models
```{r}
# Creating a dataset that contains the result of the model in a tidy way
m_control <- glm(illness ~ age + gender + income, 
                 family = "poisson", 
                 data = DoctorVisits)

summary(m_control)

m_full <- glm(illness ~ health + age + gender + income, 
              family = "poisson", 
              data = DoctorVisits)

summary(m_full)

```

# Printing the results in a table
```{r}

sjPlot::tab_model(m_control, m_full, 
                  show.aic = TRUE)

```

# Checking assumptions
```{r}
# Checking the distribution of the outcome
qplot(DoctorVisits$illness)

check_overdispersion(m_full)
check_zeroinflation(m_full)
```

# Refitting full model using negative binomial zero inflated regression

```{r}
library(pscl)

m_zinb <- zeroinfl(illness ~ health + age + gender + income, 
                   dist = "negbin", 
                   data = DoctorVisits)

sjPlot::tab_model(m_zinb,
                  show.aic = TRUE)

```



