---
title: "University admissions"
output: html_document
---

# Background

This dataset contains information on applicant chances of getting into a graduate programs. The dataset contains several predictors that may predict the chance of admission.

## Source
https://www.kaggle.com/datasets/mohansacharya/graduate-admissions

## Codebook
`codebooks/uni_admissions.txt`

# Tasks

1) Build a Beta regression model to predict admission chance.

2) Report the results. Check multicollinearity.

# Solution
```{r setup}
library(tidyverse)
library(betareg)
admissions <- read_csv(here::here("data/uni_admissions.csv"))
```

# Explore data

```{r}
# Create a histogram of the outcome
qplot(admissions$chance_of_admit)

library(GGally)
admissions %>% 
  select(gre_score:chance_of_admit) %>% 
  ggpairs()  

```


# Building the model

```{r}
admission_mod <- 
  betareg(chance_of_admit ~ gre_score + toefl_score + university_rating + 
                            sop + lor + research, 
          data = admissions)

summary(admission_mod)
tidy(admission_mod, conf.int = TRUE)

# Checking multicollinearity
performance::check_collinearity(admission_mod)

```

# Reporting results

```{r}

tab_model(admission_mod)

```

