---
title: "Marriage survival analyis"
output: html_document
---

# Background
The data on marriage dissolution in the U.S. based on a longitudinal survey. 

## Source
https://data.princeton.edu/pop509/project

## Codebook
`codebooks/sleep_exec_codebook.txt`

# Tasks

1) Fit a general model for divorce rate and report the median time for divorce.
In the next steps you need to calculate new variables from existing ones.
2) Check ethnicity difference effect where the race of the spouses (coded in `heblack` and `sheblack` variables do not match. 
3) Consider husband's education using two dummy variables:  
- When the husband has less than high school education (education less than 12 years).
- When the husband has a college education (education at least 16 years)
3) Consider the age difference between spouses. Create a model when
- the husband is older at least by 10 years
- the wife is older at least by 10 years.

# Solution
```{r setup}
library(tidyverse)
library(survival)
library(survminer)
marriage_raw <- read_csv(here::here("data/marriage.csv"))
```

# Transforming the data
```{r}
marriage <-
  marriage_raw %>% 
  mutate(mixedrace = heblack != sheblack,
         dropout = hiseduc < 12,
         college = hiseduc >= 16,
         heolder = agediff < 10,
         sheolder = agediff < -10) 
```

# Building the models

## General model (no predictors)  

```{r}
marriage_surv <- 
  survfit(Surv(time = time, event = divorced) ~ 1,
          conf.type = "log-log",
          data = marriage)

summary(marriage_surv)

ggsurvplot(marriage_surv, 
           risk.table = TRUE, 
           surv.median.line = "hv", 
           conf.int = TRUE)
```

- The median time for divorce is at: 6 years.

## Model and plot for husband's education dropout status
```{r}
dropout_surv <- 
  survfit(Surv(time = time, event = divorced) ~ dropout,
          conf.type = "log-log",
          data = marriage)

summary(dropout_surv)
survdiff(Surv(time = time, event = divorced) ~ dropout,
          data = marriage)

ggsurvplot(dropout_surv, 
           risk.table = TRUE, 
           surv.median.line = "hv", 
           conf.int = TRUE, 
           pval = TRUE)
```

## Model and plot for husband's college education
```{r}
college_surv <- 
  survfit(Surv(time = time, event = divorced) ~ college,
          conf.type = "log-log",
          data = marriage)

summary(college_surv)
survdiff(Surv(time = time, event = divorced) ~ college,
          data = marriage)

ggsurvplot(college_surv, 
           risk.table = TRUE, 
           surv.median.line = "hv", 
           conf.int = TRUE,  
           pval = TRUE)
```

## Model and plot marriage survival for age difference

Do marriages end significantly earlier when the husband is at least 10 years older?
```{r}
heolder_surv <- 
  survfit(Surv(time = time, event = divorced) ~ heolder,
          conf.type = "log-log",
          data = marriage)

summary(heolder_surv)
survdiff(Surv(time = time, event = divorced) ~ heolder,
          data = marriage)

ggsurvplot(heolder_surv, 
           risk.table = TRUE, 
           surv.median.line = "hv", 
           conf.int = TRUE, 
           pval = TRUE)
```

Do marriages end significantly earlier when the wife is at least 10 years older?
```{r}
sheolder_surv <- 
  survfit(Surv(time = time, event = divorced) ~ sheolder,
          conf.type = "log-log",
          data = marriage)

summary(sheolder_surv)
survdiff(Surv(time = time, event = divorced) ~ sheolder,
          data = marriage)

ggsurvplot(sheolder_surv, 
           risk.table = TRUE, 
           surv.median.line = "hv", 
           conf.int = TRUE, 
           pval = TRUE)
```