---
title: "DAY 4 - Furter regression types"
author: "Tamas Nagy"
date: '2022-07-13'
output: 
  html_document:
   theme: spacelab
   code_download: true
   toc: true
   toc_float: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival) # For survival analysis
library(survminer) # For survival analysis

theme_set(theme_light())
```

# Survival analysis on Game of Thrones data

Analysis question: Is there a smaller chance of survival in Game of Thrones after marriage?

We are building a Cox proportional hazards regression on got character data with the outcome variable that shows if the character is alive or not at particular ages. 

```{r}
# Read GoT character data
got_raw <- read_csv(here("data/got_characters.csv"))

# Make a few transformations to prepare data for survival analysis
got <- 
  got_raw %>% 
  drop_na(is_alive) %>% 
  mutate(death_year = if_else(is_alive == TRUE & birth_year >= 200, 301, death_year),
         status = if_else(is_alive == TRUE, 0, 1),
         age = death_year - birth_year) %>% 
  drop_na(death_year)

```

## Fit a standard model without predictors

We fit a survival model with no predictors (only time), and create the KM plot.
Median age of dying can be calculated using the `surv_median()` function.
```{r}
survival_all <- survfit(Surv(age, status) ~ 1, data = got)
ggsurvplot(survival_all, 
           surv.median.line = "hv", 
           conf.int = TRUE,
           risk.table = TRUE)

surv_median(survival_all)
```

# Check if gender and spouse predict different mortality

We are looking for main effects of marital status (if the character has a spouse) and gender.
We build three models, one with spouse, one with gender, and one with both.
The lowest AIC value is for the spouse model, so we don't include gender in the final model.
```{r}
survival_model1 <- coxph(Surv(age, status) ~ has_spouse, data = got)
survival_model2 <- coxph(Surv(age, status) ~ has_spouse + gender, data = got)
survival_model3 <- coxph(Surv(age, status) ~ gender, data = got)
    
AIC(survival_model1, survival_model2, survival_model3)

```

# Evaluate the model

We will check model summary, and the effect of having married.
We can check if there are differences between the groups using the `surv_pvalue()` function.

```{r}
# Model summary
summary(survival_model1)

# We can make a forest plot of effects
ggforest(survival_model1, data = got)

# Refit the model for visualization and further calculations
survival_model1 <- survfit(Surv(age, status) ~ has_spouse, data = got)

# Plot the two groups
ggsurvplot(survival_model1, 
           surv.median.line = "hv", 
           conf.int = TRUE,
           pval = TRUE,
           risk.table = TRUE)

surv_median(survival_model1)
surv_pvalue(survival_model1)
```


