---
title: "DAY 4 - Furter regression types"
author: "Tamas Nagy"
date: '2022-07-13'
output: 
  html_document:
   theme: spacelab
   code_download: true
   toc: true
   toc_float: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages(library(tidyverse))
library(survival) # For survival analysis
library(survminer) # For survival analysis
library(betareg) # Beta regression
library(here)
library(sjPlot)
library(broom)

theme_set(theme_light())
```

# Beta regression on proportions

We will use a dataset that contains data from households. We are trying to predict the proportion of spending on food compared to all income by income and number of persons in the household.

```{r}
# This tutorial is based on Cribari-Neto, F., & Zeileis, A. (2010). Beta Regression in R. Journal of Statistical Software, 34, 1–24. https://doi.org/10.18637/jss.v034.i02

data("FoodExpenditure", package = "betareg")
# Codebook:
# ?FoodExpenditure

# Some data cleaning
food <- 
  FoodExpenditure %>% 
  as_tibble() %>% 
  mutate(food_prop = food/income)

food
```

We build two beta regressions. In the first, the predictor will be the income, in the other, we also add the number of people in the household.

```{r}
fe_beta1 <- betareg(food_prop ~ income,     
                     data = food)

fe_beta2 <- betareg(food_prop ~ income + persons,     
                     data = food)

summary(fe_beta1)

```

## Interpreting the results

The coefficient can be interpreted as log of proportions that. After exponential transformation, the coefficients correspond to "Proportion Ratio".
For e.g. 50% food ratio means that the other 50% is spent on something else. Thus the ratio is 0.5/0.5 = 1.
A value of .99 means that one unit change in the predictor will result in a 1% decrease in the outcome. The value of 1.13 for people means that increase of one person in the household will increase the food proportion by 13%.

```{r}
tab_model(fe_beta1, fe_beta2, 
          show.aic = TRUE, 
          show.loglik = TRUE)
```

## The difference from linear regression results

Plotting the results. We can use the binomial family as the transformation that we need is the same (logit). When we compare it to a 

```{r}
food %>% 
  # Add betareg predictions to the dataset
  mutate(pred = predict(fe_beta1, .)) %>% 
  ggplot() +
  aes(x = income, y = food_prop) +
  geom_point(size = 2.5) +
  geom_line(aes(y = pred), color = "red", alpha = .5, size = 1.2) +
  # Add a linear regression line for comparison
  geom_smooth(method = lm, color = "blue", alpha = .5, lty = "dashed")


```

When the values are between mostly 0.2 and 0.8, the model is almost exactly the same as a linear regression! So the beta regression is only needed when the data is more dispersed.

# Survival analysis on Game of Thrones data

Analysis question: Is there a smaller chance of survival in Game of Thrones after marriage?

We are building a Cox proportional hazards regression on got character data with the outcome variable that shows if the character is alive or not at particular ages. 

```{r}
# Read GoT character data
got_raw <- read_csv(here("data/got_characters.csv"))

# Make a few transformations to prepare data for survival analysis
got <- 
  got_raw %>% 
  drop_na(is_alive) %>% 
  mutate(death_year = if_else(is_alive == TRUE & birth_year >= 200, 301, death_year),
         status = if_else(is_alive == TRUE, 0, 1),
         age = death_year - birth_year) %>% 
  drop_na(death_year)

```

## Fit a standard model without predictors

We fit a survival model with no predictors (only time), and create the KM plot.
Median age of dying can be calculated using the `surv_median()` function.
```{r}
survival_all <- survfit(Surv(age, status) ~ 1, data = got)

ggsurvplot(survival_all, 
           surv.median.line = "hv", 
           conf.int = TRUE,
           risk.table = TRUE)

surv_median(survival_all)
```

## Check if gender and spouse predict different mortality

We are looking for main effects of marital status (if the character has a spouse) and gender.
We build three models, one with spouse, one with gender, and one with both.
The lowest AIC value is for the spouse model, so we don't include gender in the final model.
```{r}
survival_model1 <- coxph(Surv(age, status) ~ has_spouse, data = got)
survival_model2 <- coxph(Surv(age, status) ~ has_spouse + gender, data = got)
survival_model3 <- coxph(Surv(age, status) ~ gender, data = got)
    
AIC(survival_model1, survival_model2, survival_model3)

```

## Evaluate the model

We will check model summary, and the effect of having married.
We can check if there are differences between the groups using the `surv_pvalue()` function.

```{r}
# Model summary
summary(survival_model1)

# We can make a forest plot of effects
ggforest(survival_model1, data = got)

# Refit the model for visualization and further calculations
survival_model1 <- survfit(Surv(age, status) ~ has_spouse, data = got)

# Plot the two groups
ggsurvplot(survival_model1, 
           surv.median.line = "hv", 
           conf.int = TRUE,
           pval = TRUE,
           risk.table = TRUE)

surv_median(survival_model1)
surv_pvalue(survival_model1)
```


